Staged Abstract Interpreters

### 2018-11-17 paper/30_sai.tex
revise

In the previous section, we have seen an unstaged abstract interpreter and a staged concrete interpreter, 
now we begin describing the implementation of their confluence -- a staged abstract interpreter. 
We present a principled approach to derive staged abstract interpreter
from its unstaged version. One guiding principle of our approach is the code of abstract semantics
and the code that optimizes should be separated. Thus it is also an advantage of doing staging 
for abstract interpreters: the designer of analysis has no need to rewrite the analysis, and 
the performance improvement comes almost for free, without any sacrifice of soundness or precision.
Unsurprisingly, the staged abstract interpreter we present in this section has the same abstract 
semantics with the unstaged version we presented in Section~\ref{unstaged_abs}.

\subsection{Staged Lattices} 
In Section~\ref{stagedpoly_lat}, we exploited the higher-kinded type @R@
to leave space for staging lattices, now we instantiate @R@ to @Rep@ and still use powerset as an 
example to present its staged version.

@@

The type parameter @T:Typ@ of @RepLattice@ requires the element of sets
is also can be staged, otherwise without knowing how to stage the elements in the set
we can not stage the set as well.
The methods defined operate on type @Rep[Set[T]]@, thus the underlying implementation
such as @union@ and @intersect@ will be mapped to a node in IR graph and emitted in 
the generated code. Again, other structures such as maps and tuples are also implemented
in a similar way.

\subsection{Staged Abstract Semantics} 
We have seen how to obtain a staged concrete semantics
based on types, now we take the same approach to obtain a staged abstract semantics.
\todo{what are static, what are dynamic?}
The basic operations largely kept same as the unstaged version, except the types are changed to @Rep@.
Besides, when we update the environment, the identifier @x@ is known statically, but the map of environment
has type @Rep[Map[Ident,Addr]]@, so we apply @lift@ to @x@ to make it as a next-stage value.

@@

Once more, the way we handle closures is as same as the staged concrete interpreter:
the recursive call to @ev@ with the body expression @e@ is compiled and specialized, 
the wrapper function @f@ and will be a field value of @CompiledClo@ and be generated
for the next stage. At the end, we return a singleton set.

@@

When generating code for application, we can not directly apply the callee.
Instead, we emit code that calls a next-stage function @apply_closures_norep@.
Function @apply_closures_norep@ does the same job as its unstaged version,
which non-deterministically applies multiple target closures with the argument
and latest store, and finally returns the joined value and a single store.
We provide the definition of @apply_closures_norep@ in the runtime supporting code.

@@

Our fixed-point iteration relies two caches @in@ and @out@, but the iteration 
can not be done at the current stage.
Because the @in@ and @out@ are both next-stage values, thus the test of whether 
@in@ and @out@ are equal is a generated expression to the next stage, and we can only 
know the comparison result at the next stage. 
In other words, we do not know how many iterations we need to reach the fixed-point.
To achieve this, we need to stage a function value --- @iter_aux@ is generated as a 
recursive function of type @Rep[Unit => (Value,Store)]@ and will be invoked at the next stage.

@@

However, the instrumented evaluation function that uses @in@ cache and updates @out@ cache
can be completely eliminated by staging. Each recursive call to @cached_ev@ will also be specialized
if it is applied on subexpressions of the analyzed program.

@@

Now we already obtained an end-to-end staged abstract interpreter that is able to 
specialize an analysis. However, we treat the data structures such as @Map@s as black-boxes,
which means any operations on a @Map@ becomes code in the next stage.
But, as we identified before \todo{where}, the keys of environment are identifiers in the program,
which are completely known statically. This leaves us a chance to further specialize the 
data structures. Particularly, if we are specializing a monovariant analysis, the address space 
is equivalent to the identifiers, then the environment component can be entirely eliminated, 
and the store is a partially specialized map.
\todo{maps to arrays}

@@

\subsection{Discussion}
%We have presented all of our four different interpreters as shown in Figure~\ref{confluence}.
In the literature of partial evaluation, \citeauthor{10.1007/3-540-61580-6_11} gave several
advice on what to do and not to do when specializing a concrete interpreter \cite{10.1007/3-540-61580-6_11}. 
We borrow these advice to abstract interpreters and discuss decisions we made to achieve this 
and examine some alternatives.

\paragraph{Big-step vs Small-step.}
What we implemented is a big-step, compositional \todo{or semi-compositional} abstract interpreter, 
where the "compositional" means every recursive call of our abstract interpreter is applied with proper 
substructures of the current syntactic parameters \cite{10.1007/3-540-61580-6_11}.
The compositionality ensures the specialization can be done by unfolding, as well as the termination of specialization procedure.
Nevertheless, it is also possible to specialize a small-step, operational abstract semantics --- \citeauthor{Johnson:2013:OAA:2500365.2500604}
showed this in abstract compilation \cite{Boucher:1996:ACN:647473.727587} style as one of their optimizations
of Abstract Abstract Machines \cite{Johnson:2013:OAA:2500365.2500604}. 
However, the generated abstract bytecode still requires another small-step abstract machine 
to execute, which is an additional engineering efforts.
Another alternative approach for efficient specialization is to write the abstract interpreter 
in a big-step, monadic style \cite{DBLP:journals/pacmpl/DaraisLNH17}.

@@

\paragraph{Correctness and Soundness.}
Based on the assumption that LMS preserves the semantics during staging, and indeed it is 
in our application, 
we are confident that the staged abstract interpreter does the same analysis compared 
with the unstaged one. Moreover, the optimization done by staging does not compromise
and soundness.

>>>>

In the previous sections, we have seen an unstaged abstract interpreter and a staged concrete interpreter, 
now we begin describing the implementation of their confluence -- a staged abstract interpreter. 
We present a principled approach to derive staged abstract interpreter
from its unstaged version. One guiding principle of our approach is that the code of the abstract semantics
and the code that optimizes should be separated. This it is an advantage of using staging 
for abstract interpreters: the designer of the analysis has no need to rewrite the analysis, and 
the performance improvement comes almost for free, without any sacrifice of soundness or precision.
Unsurprisingly, the staged abstract interpreter we present in this section has the same abstract 
semantics as the unstaged version we presented in Section~\ref{unstaged_abs}.

\subsection{Staged Lattices} 
In Section~\ref{stagedpoly_lat}, we exploited the higher-kinded type @R@
to leave space for staging lattices, now we instantiate the type @R@ to @Rep@ and 
still use powersets as an example to present its staged version.

@@

The type parameter @T:Typ@ of @RepLattice@ requires that the elements of sets
is can also be staged. Otherwise, without knowing how to stage the elements in the set,
we can not stage the set either.
The methods defined operate on type @Rep[Set[T]]@, thus the underlying implementation
such as @union@ and @intersect@ will be mapped to a node in the IR graph during staging and 
eventually emitted in the generated code. 
Again, other structures such as maps and tuples are implemented in a similar way.

\subsection{Staged Abstract Semantics} 
We have seen how to obtain a staged concrete semantics based on types, 
now we take the same approach to obtain a staged abstract semantics.
The basic operations are largely kept the same as in the unstaged version, except the types are changed to @Rep@.
Besides, when we update the environment, the identifier @x@ is known statically, but the environment map
has type @Rep[Map[Ident,Addr]]@, so we apply @lift@ to @x@ to turn it as a next-stage value.

@@

Once more, the way we handle closures is the same as in the staged concrete interpreter:
the recursive call to @ev@ with the body expression @e@ is compiled and specialized, 
the wrapper function @f@ will be a field value in a @CompiledClo@ object and be generated
for the next stage. At the end, we return a singleton set:

@@

When generating code for an application, we can not directly apply the callee.
Instead, we emit code that calls a next-stage function @apply_closures_norep@.
As its unstaged counterpart, function @apply_closures_norep@ 
non-deterministically applies multiple target closures with the argument
and latest store, and finally returns the joined value and a single store.
We provide the definition of @apply_closures_norep@ in the runtime supporting code.

@@

Our fixed-point iteration again relies on two caches @in@ and @out@, but the iteration 
no longer be done at the current stage.
Because the @in@ and @out@ are both next-stage values, the test of whether 
@in@ and @out@ are equal is a generated expression in the next stage, and we can only 
know the comparison result at the next stage. 
In other words, we do not know how many iterations we need to reach the fixed-point.
To achieve this, we need to stage a function value --- @iter_aux@ is generated as a 
recursive function of type @Rep[Unit => (Value,Store)]@ and will be invoked at the next stage.

@@

However, the instrumented evaluation function that uses the @in@ cache and updates the @out@ cache
can be completely eliminated by staging. Each recursive call to @cached_ev@ will also be specialized
if it is applied on subexpressions of the analyzed program.

@@

Now we have already obtained an end-to-end staged abstract interpreter that is able to 
specialize an analysis. However, we treat the data structures such as @Map@s as black-boxes,
which means any operations on a @Map@ becomes code in the next stage.
But, as we identified when introducing the generic interface, the keys of any environment maps are 
identifiers in the program, which are completely known statically. 
This leaves us a chance to further specialize the data structures. 
Assume the @Map[K,V]@ is implemented as a hash map, if the keys $K$ are known, then the indices can be computed
statically. Thus the specialized map would be an array @Array[Rep[V]]@ whose elements 
are next-stage values; all the accesses to the array is determined during staging.

Particularly, if we are specializing a monovariant analysis, the address space 
is equivalent to the identifiers, then the environment component can be entirely eliminated, 
and the store is a specialized map as array of @Rep[Value]@ elements.

@@

\subsection{Discussion}
%We have presented all of our four different interpreters as shown in Figure~\ref{confluence}.
In the literature of partial evaluation, \citeauthor{10.1007/3-540-61580-6_11} provided guidelines
on what to do and not to do when specializing a concrete interpreter \cite{10.1007/3-540-61580-6_11}. 
We borrow these guidelines and extend them to abstract interpreters. We discuss decisions we made to 
achieve this and examine some alternatives.

\paragraph{Big-step vs Small-step.}
What we implemented is a big-step, compositional \todo{or semi-compositional} abstract interpreter, 
where "compositional" means that every recursive call of our abstract interpreter is applied to proper 
substructures of the current syntactic parameters \cite{10.1007/3-540-61580-6_11}.
This compositionality ensures that specialization can be done by unfolding,
as well as that the specialization procedure terminates.
Nevertheless, it is also possible to specialize a small-step operational abstract semantics ---
\citeauthor{Johnson:2013:OAA:2500365.2500604}
showed this in abstract compilation \cite{Boucher:1996:ACN:647473.727587} style as one of their optimizations
of Abstract Abstract Machines \cite{Johnson:2013:OAA:2500365.2500604}. 
However, the generated abstract bytecode still requires another small-step abstract machine 
to execute, which is an additional engineering efforts.
Another alternative approach for efficient specialization is to write the abstract interpreter 
in a big-step, monadic style \cite{DBLP:journals/pacmpl/DaraisLNH17}.

@@

\paragraph{Correctness and Soundness.}
Based on the assumption that LMS preserves the semantics during staging,
we are confident that the staged abstract interpreter does the same analysis compared 
with the unstaged one. Moreover, the optimization done by staging does not compromise
any soundness.

### 2018-11-17 paper/50_cases.tex
revise

In Section~\ref{sai}, we have shown staging an abstract interpreter is feasible and can be 
systematic after correctly identifying the binding-times, despite the fact that the abstract 
interpreter is intended to be imprecise and easy to implement. 
In this section, we conduct several case studies to show that it is also useful and widely
applicable to different analysis.

@@

\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation (AC) as a new
implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
The idea is inspired by partial evaluation and similar with the presented paper -- the program can be known 
statically therefore the overhead of interpretation can be eliminated. 
In AC, the compiled analysis can be represented by either text or closures (higher-order functions);
the closures can be executed immediately however the textual program need to be loaded firstly.

Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a monovariant control-flow 
analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007} 
for continuation-passing style (CPS) programs. Since the analyzed program is written in CPS, the analyzer 
is essentially a big-step control-environment abstract interpreter.
Closure generation compiles the analysis as a closure which takes an environment as argument.
And the overhead of traversing the abstract syntax tree of input program also has been eliminated.

In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be 
understood and implemented as an instance of staging abstract interpreters.
We firstly revisit the original implementation of abstract compilation of 0-CFA,
and then reproduce their result by simply adding stage annotations.
The generated program of our approach improves approximately the same extent of speed,
but without changing a single line of the analyzer program (with the use of LMS).
However, closure generation requires more engineering effort, such as a whole-program 
conversion on the analyzer. Moreover, as shown in Section~\ref{staged_ds}, 
our approach is able to not only remove the interpretive overhead, but also specialize 
the data structures used in the analysis, for example, the environment that maps variables to sets of lambda.

@@

The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA for a CPS language consisting
of lambda terms, applications, @letrec@ and primtive operators. The analyses for different syntactic 
structs are decomposed to different functions, such as @analyzeCall@ and @analyzeApp@. The idea of closure
generation is to rewrite these functions, where previously they may take both static arugments and dynamic arguments,
but after the rewrite only the static arguments is taken. In this case, the static arguments are syntactic terms;
the dynamic arguments are stores. After written in AC style, the functions like @compCall@ (compiled version of @analyzeCall@) 
returns a value of type
@CompAnalysis@, i.e., a closure that takes a store and returns a store. The result of multiple calls on such functions, for example,
@compCall@ and @compArgs@ can be composed. The generated closure only takes stores, because the input program
is specialized into the closure. The code is shown in Figure~\ref{compare_ac_sai} (left).

@@

On the other side, our approach does exactly the same thing through staging: the syntactic terms are static, and 
stores are dynamic, therefore the generated code just looks-up and updates the store.
Figure~\ref{compare_ac_sai} (right) shows the code written with LMS. In fact, the only changes are
the type of @Store@ is replaced with @Rep[Store]@ indicating that the values of type @Store@  will be known
at the next stage. Indeed, a amount of additional engineering efforts are required to make this happen, 
including: the staged version @Map@ which is already included in LMS; implicit @lift@ function that transform a
current-stage constant value to next stage; next-stage representation of the syntactic terms, i.e., proper @toString@
functions of AST structs. We consider these efforts are relatively small, and does not interfere the actual analysis
we desire.

@@

The target language we presented in Section~\ref{bg_lang} is essentially a higher-order functional language.
One fundamental analysis task for functional programs is the control-flow analysis, i,e., determining
which functions will possibly be applied at each call-site. 
The abstract interpreter we used in Section~\ref{unstaged_abs} and Section~\ref{sai} is already 
a store-widened 0-CFA-like abstraction; in last section, we also reviewed AC with 0-CFA.
In this section, based on the existing staged abstract interpreter, we further
develop the staging techniques with control-flow analyses, including recoverying a more precise store model;
different polyvariant analsis, and using staging as a implementation stategy for mixed sensitivity.

@@

\subsubsection{Context/Path/Flow-Sensitivity}

\subsubsection{Mixed Sensitivity}

using different $k$ for $k$-CFA

@@

Beyond control-flow analysis, staging data structures used in the abstract interpreter, such as abstract domain.
It has been shown that specializing abstract domains with respect to the structure of analyzed program significantly 
improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
In this section, we present a similar idea for interval abstract domain using an \textit{offline} fashion by staging.

@@

Assignments, Loops, Exceptions

\subsubsection{Specializing Interval Domains}

Intervals, Numerical Analysis

Options: Data-flow analysis, taint analysis

>>>>

In Section~\ref{sai}, we have shown that staging an abstract interpreter is feasible and can be 
systematic after correctly identifying the binding-times, despite the fact that the abstract 
interpreter is intended to be imprecise and easy to implement. 
In this section, we conduct several case studies to show that this approach is also useful and
widely applicable to different analyses.

@@

\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation (AC) as a new
implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
The idea is inspired by partial evaluation similar to the present paper -- the program can be known 
statically therefore the overhead of interpretation can be eliminated. 
In AC, the compiled analysis can be represented by either text or closures (higher-order functions);
the closures can be executed immediately however the textual program need to be compiled and loaded first.

Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a monovariant control-flow 
analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007} 
for continuation-passing style (CPS) programs. Since the analyzed program is written in CPS, the analyzer 
is essentially a big-step control-environment abstract interpreter.
Closure generation compiles the analysis as a closure which takes an environment as argument.
The overhead of traversing the abstract syntax tree of input program also has been eliminated.

In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be 
understood and implemented as an instance of staging abstract interpreters.
We first revisit the original implementation of abstract compilation of 0-CFA,
and then reproduce their result by simply adding stage annotations.
The generated program of our approach improves approximately the same extent of speed,
but without changing a single line of the analyzer program (with the use of LMS).
However, closure generation requires more engineering effort, specifically a whole-program 
conversion on the analyzer. Moreover, as shown in Section~\ref{staged_ds}, 
our approach is able to not only remove the interpretive overhead, but also specialize 
the data structures used in the analysis, for example, the environment that maps variables to sets of lambda.

@@

The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA for a CPS language consisting
of lambda terms, applications, @letrec@ and primtive operators. The analyses for different syntactic 
constructs are decomposed into different functions, such as @analyzeCall@ and @analyzeApp@.
The idea of closure generation is to rewrite these functions. Where previously they may take both static
arguments and dynamic arguments, after the rewrite only the static arguments is taken.
In this case, the static arguments are syntactic terms;
the dynamic arguments are stores.
After written in AC style, functions like @compCall@ (compiled version of @analyzeCall@) return a value of type
@CompAnalysis@, i.e., a closure that takes a store and returns a store.
The result of multiple calls on such functions, for example,
@compCall@ and @compArgs@, can be composed.
The generated closure only takes stores, because the input program
is specialized into the closure. The code is shown in Figure~\ref{compare_ac_sai} (left).

@@

On the other side, our approach does exactly the same thing through staging: the syntactic terms are static, and 
stores are dynamic, therefore the generated code just looks-up and updates the store.
Figure~\ref{compare_ac_sai} (right) shows the code written with LMS. In fact, the only changes are
the type of @Store@ is replaced with @Rep[Store]@ indicating that the values of type @Store@  will be known
at the next stage. Indeed, additional engineering efforts are required to make this happen, 
including: the staged version @Map@ which is already included in LMS; implicit @lift@ function that transform a
current-stage constant value to next stage; next-stage representation of the syntactic terms, i.e., proper @toString@
functions of AST structs. We consider these efforts relatively small, and they do not interfere the actual analysis
we desire.

@@

The target language we presented in Section~\ref{bg_lang} is essentially a higher-order functional language.
One fundamental analysis task for functional programs is control-flow analysis, i,e., determining
which functions will possibly be applied at each call-site. 
The abstract interpreter we used in Section~\ref{unstaged_abs} and Section~\ref{sai} is a store-widened 
0-CFA-like abstraction, moreover it is also a pushdown control-flow analysis; 
in the last section, we also reviewed AC with finite-state 0-CFA.
In this section, based on the existing staged abstract interpreter, we further
develop the staging techniques with control-flow analyses, including recoverying a more precise store model and
adding context-sensitivity to the analysis.
%, and using staging as an implementation stategy for mixed sensitivity.

@@

The store-widened analysis we implemented in Section~\ref{sai} uses a single store to approximate
the runtime store. It can be efficently computed in polynomial time together with 0-CFA-like abstraction, 
but sometimes we may desire a 
more precise result that distinguishes the final values and stores for different closure targets. 
To achieve this goal,
we need to tweak our abstract interpreter and type instantiation. The answer type @Ans@ is changed to
a set of @VS@s where a @VS@ is a pair of sets of abstract values (such as closures or abstract numbers)
and a store.

\begin{lstlisting}
type VS = (Set[AbsValue], Store)
type Ans = R[Set[VS]]
\end{lstlisting}

Note that the type @Ans@ uses our stage polymorphic type @R@, meaning that under staging the type @Ans@
represents a next stage value. Then, our generic interpreter is also changed when handling function 
applications.

\begin{lstlisting}
case App(e1, e2) =>
  val e1ans = ev(e1, $\rho$, $\sigma$)
  val e1vs  = choices(e1ans)
  val e2ans = ev(e2, $\rho$, e1vs.$\sigma$)
  val e2vs  = choices(e2ans)
  apply_closure(ev)(e1vs.v, e2vs.v, e2vs.$\sigma$)
\end{lstlisting}

The idea to handle the application is to explore all possible closures from @e1@ and meanwhile
use the latest store.
What @choices@ does is similar to McCarthy's @amb@ operator ~\cite{MCCARTHY196333}: it non-deterministically returns an
element of type @VS@ from its argument, e.g., @e1ans@, to its right-hand side receiver, i.e., @e1vs@.
The function @choices@ internally uses the delimited control operator @shift@ to capture the continuation,
which is the code block after its call site. 
This allows us to perform nesting depth-first evaluation for function application while still writing 
the program in direct-style \cite{Wei:2018:RAA:3243631.3236800}. 
Again, we can stage this part as we did for the naive abstract interpreter.

\subsubsection{Context-Sensitivity}

We add k-CFA-like context-sensitivity to the analysis by introducing an abstract timestamp, 
whose concrete instantiation is a finite list of expressions that track $k$ recent calling contexts.
The definition of abstract addresses is changed to a tuple of identifiers and the time it get allocated,
meaning that this address points to some values under such calling context.
If $k$ is 0, we obtain a monovariant analysis as demonstrated before; if $k > 0$, we obtain
a family of analysis with increasing precision.

\begin{lstlisting}
type Time = List[Expr]
\end{lstlisting}

Every time when we call the @eval@ function, we refresh the timestamp by calling function @tick@,
which returns a new timestamp. Here we adopt a $k$-CFA-like allocation strategy, therefore
the @tick@ function can be implemented as appending the current expression being evaluated to the existing
calling context, and then taking the first $k$ elements from the list.

\begin{lstlisting}
def tick(e: Expr, $\tau$: R[Time]): R[Time] = (e :: $\tau$).takewhile
def eval(ev: EvalFun)
  (e: Expr, $\rho$: R[Env], $\sigma$: R[Store], $\tau$: R[Time]): Ans = {
    val $\tau$_* = tick(e, $\tau$)
    e match {
      case Lit(i) => ...
      ...
    }
  }
\end{lstlisting}

Accordingly, the type of return value is accompanied by the timestamp. For a recursive call of @ev@,
which would also return the latest timestamp, and that timestamp will be used for the evaluation afterward.

\begin{lstlisting}
type VST = (Value, Store, Time)
type Ans = R[Set[VST]]
\end{lstlisting}

Using other allocation strategies to achieve different sensitivities is also possible \cite{DBLP:conf/icfp/Gilray0M16} 
and can be staged under our framework.

%\subsubsection{Mixed Sensitivity}

%using different $k$ for $k$-CFA

@@

Now we consider in a first-order imperative language, we may care more about the data-flow because the control-flow
is relatively easy to obtain. In this section, we show the staging of other abstract domains, particularly an 
interval domain for numbers.
It has been shown that specializing abstract domains with respect to the structure of analyzed program significantly 
improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
In this section, we first show how to support imperative language features in the generic abstract interpreter.
Then we present a similar idea for the interval domain and show that the specialization is feasible by staging systematically.

@@

To evaluates an assignment, we first evaluate its right-hand side, and then put the value
into the slot where the address of @v@ points to.
For simplicity, we elect to make the value of the assignment be an @void()@ value.

\begin{lstlisting}
case Assign(x, e) =>
  val (v, $\sigma$_*) = ev(e, $\rho$, $\sigma$)
  (void(), put($\sigma$_*, get($\rho$, x), v))
\end{lstlisting}

To evaluates a @while@ loop statement, we evaluate the condition first.
Then similar to how we treat @branch0@, we have a generic @branch@ function but works on boolean
values. For the @true@ branch, we recursively call @ev@ on a newly constructed expression
@Seq(e, While(t, e))@ meaning that first evaluates @e@ and the repeat the loop.
Otherwise, for the other branch, we simply return a void value and current store.

\begin{lstlisting}
case While(t, e) =>
  val (tv, t$\sigma$) = ev(t, $\rho$, $\sigma$)
  branch(tv, ev(Seq(e, While(t, e)), $\rho$, t$\sigma$), (void(), t$\sigma$))
\end{lstlisting}

\subsubsection{Staged Interval}

\begin{lstlisting}
case class Interval(lb: Rep[Double], ub: Rep[Double]) {
  def +(that: Interval): Interval = that match {
    case Interval(lb_, ub_) => Interval(lb + lb_, ub + ub_)
  }
  def -(that: Interval): Interval = that match {
    case Interval(lb_, ub_) => Interval(lb - lb_, ub_ - ub) 
  }
  ...
}
\end{lstlisting}

### 2018-11-17 paper/70_related.tex
revise

\paragraph{Abstract Interpreters}
Abstract interpretation was proposed as a semantic-based approach to construct sound static analysis by approximation \cite{DBLP:conf/popl/CousotC77}.
As semantic artifacts, the Abstracting Abstract Machines (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
approach shows that abstract interpreters can be derived systematically from concrete semantic artifacts.
The big-step abstract interpreters we presented in this paper are also inspired by the AAM framework,
which are later adopted for big-step abstract interpreters \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}.
\citet{DBLP:conf/cc/CousotC02} also proposed an abstract interpretation framework for modular analysis.
\citet{DBLP:conf/popl/CalcagnoDOY09} developed compositional shape analysis by using bi-abduction.
The big-step abstract interpreter we presented in this paper is inspired by \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}.

\textbf{Meta-Programming}.

Partial evaluation. 
Reflexive language \cite{DBLP:conf/gpce/Asai14}
Collapsing Towers of Interpreters \cite{Amin:2017:CTI:3177123.3158140}

The Lightweight Modular Staging framework in Scala
Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 
and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
Compared with the LMS approach in Scala, MetaML/MetaOCaml use term-level annotations such as brackets, 
escape, and run. Notwithstanding, we use LMS in this paper, the idea of staging an abstract interpreter
still applies with other MSP implementations.

>>>>

\paragraph{Abstract Interpreters}
Abstract interpretation was proposed as a semantic-based approach to construct sound static analysis by approximation \cite{DBLP:conf/popl/CousotC77}.
As semantic artifacts, the Abstracting Abstract Machines (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
approach shows that abstract interpreters can be derived systematically from concrete semantic artifacts.
The big-step abstract interpreters we presented in this paper are also inspired by the AAM framework,
which are later adopted for big-step abstract interpreters \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}.
\citet{DBLP:conf/cc/CousotC02} also proposed an abstract interpretation framework for modular analysis.
\citet{DBLP:conf/popl/CalcagnoDOY09} developed compositional shape analysis by using bi-abduction.
The big-step abstract interpreter we presented in this paper is inspired by \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}. 
\citet{Keidel:2018:CSP:3243631.3236767} demonstrate a similar implementation of definitional abstract interpreter using arrows.

\paragraph{Meta-Programming}
The Lightweight Modular Staging framework relies on types as stage annotations, and the staging
procedure is modular.
Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 
and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
Compared with the LMS approach in Scala, MetaML/MetaOCaml use term-level annotations such as brackets, 
escape, and run. Notwithstanding, we use LMS in this paper, the idea of staging an abstract interpreter
still applies with other MSP implementations.
Partial evaluation as an automatic technique is studied comprehensively
\cite{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.
Futamura projections reveals close relations between interpreters and compilers \cite{Futamura1999, futamura1971partial}.
\citet{DBLP:conf/gpce/Asai14} studies the compilation of reflective language using MSP and MetaOCaml.
\citeauthor{Amin:2017:CTI:3177123.3158140} shows how to collapse a tower of concrete interpreters by MSP and LMS framework.
